---
title: "Ukraine War: Russian Losses Timeline"
author: "Cormac Nolan"
date: Sys.Date()
output: html_document
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
---

```{r setup, include=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(forecast)
library(here)
```

## Import

Data sourced from the [Kyiv Independent Twitter feed](https://twitter.com/KyivIndependent/media) who sourced it from the Ukrainian Defense Ministry.

These are imperfect estimates, given during chaotic wartime conditions, from an institution that is would be incentivised to inflate its figures. This isn't to say that the Defense Ministry is consciously inflating them, but it bears remembering when looking at this data.

```{r import}
source(here("data/ukraine_defense_ministry_russian_loss_estimates.R"))

campaign_length = as.integer(max(russian_losses$report_date) - min(russian_losses$report_date))
campaign_length_weeks = ceiling(campaign_length / 7)

russian_losses
 
```

## Rates

First we need to apply interpolation in case there were any missed days, linear is fine.

```{r, fig.height=7}

russian_losses_all_dates <-
  dplyr::tibble(report_date = min(russian_losses$report_date) + 0:campaign_length) %>%
  left_join(russian_losses, by="report_date")


russian_losses_all_dates <- 
  russian_losses_all_dates %>% 
  mutate(across(where(~is.numeric(.x)), .fns=forecast::na.interp)) %>%
  mutate(across(where(~is.numeric(.x)), .fns=as.numeric)) %>% 
  mutate(across(where(~is.numeric(.x)), .fns=round))


russian_losses_all_dates <- 
  russian_losses_all_dates %>% 
  mutate(campaign_week_number = c(0, rep(1:campaign_length_weeks, each=7)[1:(campaign_length)]))
  
russian_losses_all_dates$report_date
```

Now let's calculate rates...

```{r}
daily_diff <- function(val){
  val - dplyr::lag(val) 
}

calculate_loss_rate <- function(df){
  
  df %>% 
    mutate(across(where(is.numeric), .fns=daily_diff), .keep = "unused") %>% 
    tidyr::drop_na()
}


russian_losses_rates <- 
  russian_losses_all_dates %>% 
    select(-campaign_week_number) %>% 
    calculate_loss_rate()


select(russian_losses_rates, report_date, soldiers, tanks) %>% 
  head(5)
```

### Important Events

```{r}
source(here("./data/ukraine_war_important_events.R"))
important_events %>%  head(1)
```

## Rate Plots Timeline

#### APC's

```{r}

ukraine_blue <- "#0057b7"
ukraine_blue_light <- "#ACC8E7"
ukraine_blue_very_light <- "#EDF4FA"
ukraine_blue_dark <- "#00346d"
ukraine_yellow <- "#ffd700"
ukraine_yellow_darkened <- "#d6b500"
text_colour <- "#555555"

ukraine_plot_theme <- function(){
  theme(plot.title = element_text(size = 20, colour = text_colour),
      strip.text.x = element_text(size = 16, colour = text_colour),
      axis.title = element_text(size = 15, face="bold", colour = text_colour),
      axis.text.x = element_text(
        angle = 30, hjust = 1, colour = text_colour, size = 13
        ),
      axis.text.y = element_text(size = 13),
      panel.background = element_rect(
        fill = ukraine_blue_very_light, colour = ukraine_blue_light
        )
      )
}

format_names <- function(in_string){
  out_string <- in_string %>% 
    stringr::str_replace_all("_", " ") %>% 
    stringr::str_to_title()
  
  out_string <- ifelse(out_string == "Uav", yes = "UAV", no = out_string)
  out_string
}

prepare_rate_plot_data <- function(rates_df, time_col, ...) {
  rates_df %>% 
  select({{time_col}}, ...) %>% 
    rename_with(.fn=format_names, .cols=-contains("date")) %>% 
    tidyr::pivot_longer(
      -{{time_col}}, names_to = "category", values_to = "loss_count"
      )
  
}


rate_plot <- function(
    rates_df, events_df, category_for_plot, plot_title, plot_x, plot_y
  ) {
  
  current_plot_rates <-   rates_df %>% 
    filter(category == category_for_plot)
  
  current_plot_rates %>% 
    ggplot(aes(report_date, loss_count)) +
    geom_bar(stat = "identity", alpha = 0.5, fill = ukraine_yellow_darkened) +
    geom_smooth(
      se=FALSE, 
      span=0.7,
      method = "loess", 
      formula = y ~ x, 
      colour = ukraine_blue, 
      size = 1.5
    ) +
    geom_vline(mapping = aes(xintercept = report_date), 
               data = events_df, 
               colour = ukraine_blue_dark,
               linetype="dashed", alpha = 0.8, size = 1.1) +
    geom_text(
      mapping = aes(
        x = report_date + 1,
        label = event_name,
        y = max(current_plot_rates$loss_count) / 2
      ),
      data = mutate(events_df, category = category_for_plot),
      colour = ukraine_blue_dark,
      angle = 90,
      size = 5
    ) +
    theme_minimal() +
    ukraine_plot_theme() +
    labs(title = plot_title, x = plot_x,y = plot_y)
}


```

```{r, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
russian_losses_rates %>% 
  prepare_rate_plot_data(
    report_date, armoured_personel_carriers
  ) %>% 
  rate_plot(
    events_df = filter(important_events, event_category == "land"),
    category_for_plot = "Armoured Personel Carriers",
    plot_title = "Russian Daily Loss Rates: APC's", 
    plot_x = "Report Date", 
    plot_y="Loss Count"
)
```

#### Aircraft

```{r, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
russian_losses_rates %>% 
  prepare_rate_plot_data(
    report_date, uav
  ) %>% 
  rate_plot(
    events_df = filter(important_events, event_category == "land"),
    category_for_plot = "UAV",
    plot_title = "Russian Daily Loss Rates: UAV's", 
    plot_x = "Report Date", 
    plot_y="Loss Count"
)

```

```{r, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
russian_losses_rates %>% 
  prepare_rate_plot_data(
    report_date, helicopters
  ) %>% 
  rate_plot(
    events_df = filter(important_events, event_category == "land"),
    category_for_plot = "Helicopters",
    plot_title = "Russian Daily Loss Rates: Helicopters", 
    plot_x = "Report Date", 
    plot_y="Loss Count"
)

```

#### Artillery

```{r, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
russian_losses_rates %>% 
  prepare_rate_plot_data(
    report_date, artillery_pieces
  ) %>% 
  rate_plot(
    events_df = filter(important_events, event_category == "land"),
    category_for_plot = "Artillery Pieces",
    plot_title = "Russian Daily Loss Rates: Artillery", 
    plot_x = "Report Date", 
    plot_y="Loss Count"
)
```
